Ethnicity Sensitive Author Disambiguation Using Semi-supervised Learning

### 2016-06-28 paper/author_disambiguation.tex
Shorter version of Hussein's affiliation
blocking and meta-data + small and compact title section

\usepackage{xspace}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{cite}
\usepackage[hidelinks]{hyperref}
\newcommand*{\eg}{e.g.\@\xspace}
\newcommand*{\ie}{i.e.\@\xspace}
\usepackage{color}
\usepackage{microtype}

@@

\institute{CERN, Switzerland,\\
\email{\{g.louppe, eamonn.james.maguire\}@cern.ch},
\and
Univ Lyon, ISH Lyon, USR 3385, CNRS, and ERIC lab., EA 3083, Lyon 2, France,\\
\email{hussein.al-natsheh@ish-lyon.cnrs.fr}
\and
University of Warsaw, Poland,\\
\email{msusik@student.uw.edu.pl}
}

@@

Solutions to author disambiguation have been proposed from various
communities \cite{liu2014author}. On the one hand, libraries have maintained
authorship control through manual curation, either in a centralized way by
hiring professional collaborators or through developing services that invite authors to register
their publications themselves (\eg, Google Scholar or Inspire-HEP).
Recent efforts to create persistent digital identifiers assigned to researchers (\eg, ORCID or ResearcherID),
with the objective to embed these identifiers in the submission workflow of publishers
or repositories (\eg, Elsevier, arXiv or Inspire-HEP), would univocally solve any disambiguation issue.
As the centralized manual authorship control is expensive and the success of persistent digital
identifiers requires large and ubiquitous adoption by both researchers and publishers, fully
automated machine learning-based methods have been proposed to
provide immediate, less costly, and satisfactory solutions to author
disambiguation. In this work, we study how labeled data obtained through manual curation (either centralized or
crowdsourced) can be exploited (i) to learn an accurate classifier for
identifying coreferring authors, and (ii) to guide the clustering of scientific
publications by distinct authors in a semi-supervised way.
Our analysis of parameters and features of this large dataset reveal that the general pipeline
commonly used in existing solutions is an effective approach for author disambiguation.
Moreover, we propose better strategies for blocking based on the
phonetization of author names to increase recall and ethnicity-sensitive features for learning a
linkage function which tailor our author disambiguation to non-Western author names.

@@

As in previous works, the first part of our algorithm consists of dividing
signatures ${\cal S}$ into disjoint subsets ${\cal S}_{b_0}, ..., {\cal
S}_{b_{K-1}}$, or \textit{blocks} \cite{fellegi69}, followed by carrying out
author disambiguation on each one of these blocks independently.
By doing so, the computational complexity of clustering (see Section~\ref{methods:clustering})
typically reduces from $O(|{\cal S}|^2)$ to $O(\sum_b |{\cal S}_b|^2)$.
Since disambiguation is performed independently per block, a good blocking strategy should be
designed such that signatures from the same author are all mapped to the same
block, otherwise their correct clustering would not be possible in later stages of the workflow.
As a result, blocking should be a balance between reduced complexity and maximum recall.

@@

\subsubsection{Input features.} Following previous works, pairs of signatures $(s_1, s_2)$ are first transformed to vectors $v \in \mathbb{R}^p$
by building so-called similarity profiles \cite{treeratpituk2009disambiguating} on which supervised learning is carried out.
In this work, we design and evaluate fifteen standard input
features \cite{ferreira2012brief, levin2012citation} based on the comparison of signature fields, as reported in the first
half of Table~\ref{table:features}.
As an illustrative example, the \textit{Full name} feature corresponds to the similarity between the (full)
author name fields of the two signatures, as measured using as combination
operator the cosine similarity between their respective $(n,m)$-\emph{TF-IDF} vector
representations\footnote{$(n,m)-$\emph{TF-IDF} vectors are \emph{TF-IDF} vectors computed
from $n$, $n+1$, ..., $m$-grams.}.

>>>>

\usepackage{xspace}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{cite}
\usepackage[hidelinks]{hyperref}
\newcommand*{\eg}{e.g.\@\xspace}
\newcommand*{\ie}{i.e.\@\xspace}
\usepackage{color}
\usepackage{microtype}
\usepackage[small,compact]{titlesec}

@@

\institute{CERN, Switzerland,\\
\email{\{g.louppe, eamonn.james.maguire\}@cern.ch},
\and
ISH, USR 3385, CNRS, and ERIC lab, EA 3083, Univ Lyon 2, France,\\
\email{hussein.al-natsheh@ish-lyon.cnrs.fr}
\and
University of Warsaw, Poland,\\
\email{msusik@student.uw.edu.pl}
}

@@

Solutions to author disambiguation have been proposed from various
communities \cite{liu2014author}. On the one hand, libraries have maintained
authorship control through manual curation, either in a centralized way by
hiring professional collaborators or through developing services that invite authors to register
their publications themselves (\eg, Google Scholar or Inspire-HEP).
Recent efforts to create persistent digital identifiers assigned to researchers (\eg, ORCID or ResearcherID),
with the objective to embed these identifiers in the submission workflow of publishers
or repositories (\eg, Elsevier or arXiv), would univocally solve any disambiguation issue.
As the centralized manual authorship control is expensive and the success of persistent digital
identifiers requires large and ubiquitous adoption by both researchers and publishers, fully
automated machine learning-based methods have been proposed to
provide immediate, less costly, and satisfactory solutions to author
disambiguation. In this work, we study how labeled data obtained through manual curation (either centralized or
crowdsourced) can be exploited (i) to learn an accurate classifier for
identifying coreferring authors, and (ii) to guide the clustering of scientific
publications by distinct authors in a semi-supervised way.
Our analysis of parameters and features of this large dataset reveal that the general pipeline
commonly used in existing solutions is an effective approach for author disambiguation.
Moreover, we propose better strategies for blocking (\ie, partitioning) based on the
phonetization of author names to increase recall and ethnicity-sensitive features for learning a
linkage function which tailor our author disambiguation to non-Western author names.

@@

As in previous works, the first part of our algorithm consists of dividing
signatures ${\cal S}$ into disjoint subsets ${\cal S}_{b_0}, ..., {\cal
S}_{b_{K-1}}$, or \textit{blocks} (\ie partitions) \cite{fellegi69}, followed by carrying out
author disambiguation on each one of these blocks independently.
By doing so, the computational complexity of clustering (see Section~\ref{methods:clustering})
typically reduces from $O(|{\cal S}|^2)$ to $O(\sum_b |{\cal S}_b|^2)$.
Since disambiguation is performed independently per block, a good blocking strategy should be
designed such that signatures from the same author are all mapped to the same
block, otherwise their correct clustering would not be possible in later stages of the workflow.
As a result, blocking should be a balance between reduced complexity and maximum recall.

@@

\subsubsection{Input features.} Following previous works, pairs of signatures $(s_1, s_2)$ are first transformed to vectors $v \in \mathbb{R}^p$
by building so-called similarity profiles \cite{treeratpituk2009disambiguating} on which supervised learning is carried out.
In this work, we design and evaluate fifteen standard input
features \cite{ferreira2012brief, levin2012citation} based on the comparison of signature fields, as reported in the first
half of Table~\ref{table:features}. Noteworthy, the author metadata, both provided or derived, are far more important than the publication content itself.
As an illustrative example, the \textit{Full name} feature corresponds to the similarity between the (full)
author name fields of the two signatures, as measured using as combination
operator the cosine similarity between their respective $(n,m)$-\emph{TF-IDF} vector
representations\footnote{$(n,m)-$\emph{TF-IDF} vectors are \emph{TF-IDF} vectors computed
from $n$, $n+1$, ..., $m$-grams.}.
